Contracts for unions and intersections are hard for Nickel

### 2021-06-03 paper.tex
Fix comma at beginning of line
This is a bit brutal to be honest… and may be a bad idea.
Wip: Arnaud's pass through the introduction
Very not finished but I'll probably only be done tomorrow.

Build systems, package managers, operating systems, cloud infrastructure,
continuous integration and web services\resolved{Also continuous integration}
are examples of modern complex software systems which require an extensive and
non-trivial configuration. To handle this complexity, the declarative approach
has become more and more prominent: instead of applying a sequence of
transformations to a system in order to reach a desired state, this state is
described as static textual data. Some tool is then in charge of applying the
necessary transformations. This approach makes the configuration persistent,
reproducible, versionable and subject to static analysis. Text-based
configurations are thus led to play an important role in critical aspects of
software engineering, including security, availability, and maintainability.

However, static text-based configurations show their limits beyond a certain
size and complexity. Data cannot be transformed, combined, shared, nor validated
(without resorting to separate schemas and tools).  This lack of expressivity
results at best in boilerplate and duplication of information, but it can also
make data invalid or inconsistent.  Correctly configuring a modern system is hard
and failures may have substantial negative consequences.

@@

Figure~\ref{fig:contract-reporting-wo} and Figure~\ref{fig:contract-reporting-w}
give an example of a situation where adding a simple contract annotation
\nickel{Str -> Str} to a function greatly improves the error message. In the
original program of Figure~\ref{fig:contract-reporting-wo}, the
\nickel{catHosts} function takes a string argument, append it to a list that is
then folded. However, the programmer forgot one parameter when calling to
\nickel{makeHost}, thus passing a partially applied function instead of a string
to \nickel{catHosts}. Because this argument is put in a list and folded, the
error is raised later on and points to inside the code of \nickel{catHosts} and
\nickel{makeHost}, which may be totally unknown to the caller.

@@

% LocalWords:  Felleisen

>>>>

A configuration language is a language concerned with describing the
configuration of an application. In traditional configuration
languages, such as YAML, TOML, or JSON, the configuration is fully,
and explicitly, spelt out.

However, with the advent of DevOps, configurations have been extended
to describe the entire state of a computer, or even a fleet of
computers. For instance, with Kubernetes you need to configure a large
fleet of (possibly replicated) docker containers. To describe this
sort of configurations, you really want to be able to re-use and
abstract parts of the configuration, like traditional programming
languages do. To meet this need, languages such as Cue~\cite{?},
Dhall~\cite{?}, Jsonnet~\cite{?}, or Nickel~\cite{NickelRepo}, where
configurations are generated rather than spelt out, were created.

Another example is continuous integration systems: it's fairly typical
to need a matrix of job, wherein the same tests are run on different
infrastructures, or with different versions of a compiler. Traditional
configuration would have you write the same steps for each
infrastructure. It's much better to write the steps once, and
instantiate them for each infrastructure. Continuous integration
systems typically do this using a templating system layered on top of
YAML. Each of the configuration-generating languages above allow such
job-matrix definition natively.

\subsection*{Nickel}

In this article, we will use the Nickel language~\cite{NickelRepo} as
illustration and motivation. At its core, Nickel is the JSON data
model, that is dictionaries (written \nickel{\{field1 = value1, ...,
  fieldn = valuen\}}) and arrays (\nickel{[x1, ..., xn]}), augmented
with functions (\nickel{fun arg1 ... argn => body}) and
let-definitions (\nickel{let id = value in exp}).\unsure{Arnaud: this
  paragraph looks kind of bad, maybe it'll be better to put each
  syntax entry in an enumeration, we will see if there is still room
  for this}

A Nickel configuration is then evaluated to an explicit configuration,
\emph{e.g.}  in JSON, which can then be consumed by an
application. Therefore a design constraint of Nickel is any Nickel
data must have a straightforward interpretation in JSON.

A useful feature of a configuration language is to provide facilities
for schema validation. That is, help answer questions like: does our
configuration have all the required fields? does this field indeed
contains a URL?

These are an inherently dynamic question, as they are all questions
about the evaluated configuration.

\unsure{Arnaud: What's below this point in the introduction is being reworked}

@@

Figure~\ref{fig:contract-reporting-wo} and Figure~\ref{fig:contract-reporting-w}
give an example of a situation where adding a simple contract annotation
\nickel{Str -> Str} to a function greatly improves the error message. In the
original program of Figure~\ref{fig:contract-reporting-wo}, the
\nickel{catHosts} function takes a string argument, append it to a list that is
then folded. However, the programmer forgot one parameter when calling to
\hbox{\nickel{makeHost},} thus passing a partially applied function instead of a string
to \nickel{catHosts}. Because this argument is put in a list and folded, the
error is raised later on and points to inside the code of \nickel{catHosts} and
\nickel{makeHost}, which may be totally unknown to the caller.

@@

% LocalWords:  Felleisen JSON TOML YAML templating Jsonnet Dhall
% LocalWords:  Kubernetes DevOps natively
